RadLib: a radiative heat transfer model library for CFD

### 2021-03-24 submissions/CPC_response_to_reviewers.md

## Reviewer 1

This manuscript introduces “RadLib,” a C++ library to calculate absorption coefficients for use in radiative heat transfer calculations. The library has reasonable value to the research community, despite a few limitations in its scope and design. The library is self-contained (doesn’t have any external dependencies), which is convenient.

I have a number of comments on the manuscript as well as the software itself which the authors should consider to improve the manuscript and software.

### Manuscript:
1.There is no general user documentation provided in section 3. Figure 1 includes information that is more algorithmic than illustrative of the API.

2. Section 5 makes several editorial claims that are quite subjective, and not well-supported:
	1. First sentence: “historically difficult to implement … complexity and high computational costs”
		- Property evaluation is not the expensive part of a radiation calculation. Solving the RTE is the expensive part.
        - I wouldn’t characterize the property evaluation as particularly complex in comparison with other aspects of combustion simulations (kinetics, thermodynamics, transport properties).

    1. At the end of the first paragraph, the authors seem to imply that RadLib will solve the challenges of radiation modeling. Again, it is the RTE solver, not the property evaluation, that is the largest time-sink in terms of computer and human resources.

    1. These kinds of statements persist throughout the second paragraph as well. There are limitations in RadLib such as lack of scattering treatment for particle-laden flows, inability to include more species, lack of any line-by-line capabilities, etc. The library checks a few useful boxes, but the verbiage here seems to take more credit than the authors may be due.

3. Section 5 claims that “RadLib’s modular framework is designed to easily accommodate new models as well,” which is a bit of an overstatement. See comments below w.r.t. assumptions on species considered, hard-coded polynomial coefficients, etc. all of which reduce extensibility.

**In order to avoid subjective language, this sentence has been revised as follows: "RadLib's modular framework can also accomodate new models, . . .". Specific extensibility issues are addressed below with their respective reviewer comments.**

4.There are several minor issues that need to be addressed:

    1. Abstract: “well-written” is subjective and should be removed.

	** The offending sentence (actually located in the "Solution method" section of the Program Summary) has been revised to eliminate subjective language.**

    1. Introduction: second sentence: “Combustion simulations…” is a run-on sentence.

	** Technically, no, this is not a run-on sentence, but its structure is awkward and inhibits readability. It has been revised to the following: "For example, combustion simulations are often complicated by the need for accurate radiation modeling, which can be difficult to implement and computationally expensive relative to overall simulation cost."**

    1. Page 10: “[INSERT RUNTIME COMPARISONS HERE]”

    3. Section 7 heading: “Interes”

	**Fixed section heading. Now correctly reads "Declaration of Competing Interests".**

### Software deficiencies & suggestions:

1. API doxygen documentation is non-standard. It appears that the doxygen documentation is provided in the .cc files rather than the header files, which is where it is typically found. Downstream developers would typically be working from headers and the installed library, so it is convention to have doxygen documentation in the header only. This also reduces clutter in the implementation files.

** The authors were previously unaware of any such standards or conventions, and follow up research indicates that this is a stylistic preference rather than a flaw in the software. The locations of doxygen documentation blocks were chosen for overall readability as they tended to clutter the header files and were less obtrusive when located in the implementation files. However, the reviewer does make a valid point regarding usage by downstream developers, and while we would expect the typical RadLib user to be comfortable working with both header files and source code, it is worthwhile to consider all possible usages. Because this is an issue of style rather than function, no immediate changes have been made to the code or its documentation, but the authors appreciate this information about standard practice and will consider its implications moving forward.**

2. make_examples.sh:
    1.	will only work with a g++ compiler in the user’s path.
    1.	Is referred to as build_examples.sh in the README.md file
    1.	is redundant with the cmake build system and isn’t portable. It should probably be removed.
3. It appears that there are no tests associated with the library. This is a bit surprising - I would expect regression test coverage on the basic API functionality.
4. CMake build system:
    1.	Consider installing a “RadLib.cmake” file for downstream usage by CMake-based projects. This helps downstream build systems configure for RadLib usage (setting include paths, etc.).
    1.	probably shouldn’t specify optimize or debug flags directly, as CMake will provide appropriate values on most platforms.
    1.	the default build type isn’t defaulting to Release for me; it remains blank.
    1.	I suggest not using verbose makefiles by default.
    1.	Using GLOB to install the headers is not the appropriate CMake approach.
    1.	the data files are not copied into the examples build directory, meaning that the example executables do not run.
    1.	The run-examples.sh script is not helpful as it is not installed into the build directory where the executables are produced and doesn’t have path information to actually run the executables.
    1.	Using “doxygen” as a target is not ideal since it conflicts with the “doxygen” executable name. It still works fine though...

5. It appears that line-by-line data from HITEMP/HITRAN files cannot be loaded into RadLib. That’s an unfortunate limitation, but not a deal-breaker.

** At present, RadLib's models cannot accommodate such data input, but this is a limitation of the models themselves, not the software. The appropriate line-by-line data is included within the RadLib package, which the authors consider a convenience rather than a hindrance. The ability to load database files directly into RadLib could be useful to researchers looking to extend RadLib's model library, however. The authors appreciate the suggestion and will consider it for future development.**

6. It looks like the software hard-codes Plank mean absorption coefficients as (fourth-order?) polynomial fits? Same for WSGG. There is no discussion of why this was chosen, how the coefficients were determined, or a characterization of its accuracy. Do you observe any problematic behavior with the fourth order polynomials over temperatures from 300 to 3000 K?  Are these interpolants or regression of data?

** This discussion was purposely omitted because it does not align with the goal of the paper, which is to present RadLib as a software tool for calculating radiation property data, not to evaluate or compare the included models. All of the reviewer's questions are addressed in detail in the referenced literature (Planck Mean coefficients in [3,5] and WSGG in [10,11]). The authors believe that this is an issue of scope rather than missing information, but they are willing to extend the discussion at the editor's discretion. **

7. The two previous items combined lead to a significant limitation of RadLib: one cannot easily add new species or modify existing data that is used to generate the absorption coefficients since there has been internal preprocessing of (an unknown subset of) the HITEMP/HITRAN databases to produce curve fits. And given that this process is opaque, it isn’t reproducible.

8. RadLib only considers CO2 and H2O for WSGG, which is fairly limited. Even for Plank mean, only H2O, CO2, CO and CH4 are considered. NO and OH would be nice to include. It appears that including additional species would be a non-trivial undertaking.

** All WSGG correlations consider only CO2-H2O mixtures with the exception of specialized subset of WSGG models that consider CO2-H2O-soot mixtures. This WSGG model was chosen for RadLib due to its popularity, relative simplicity, and good performance. Note that WSGG models usually address radiation in combustion environments. RadLib implements the WSGG model presented by Bordbar et al. in references [10] and [11], which was developed specifically for use in oxy-fired combustion simulations. CO2 and H2O are the major products of all combustion processes, and CO is a major product of rich combustion. NO and OH are minor species that, while important to other aspects of combustion processes, do not typically participate significantly in radiative heat transfer processes. For more information, please refer to Section 20.6 of reference [2] and/or Section 3.4.1 of *Radiative Heat Transfer in Turbulent Combustion Systems* by Modest and Haworth.

It should also be noted that the Planck Mean absorption coefficients model as implemented in RadLib refers specifically to the radiation model in [3,5], which is one of the most common radiation models applied to turbulent combustion simulations. This is why Section 2.1 explicitly references the TNF radiation model and its calculated Planck Mean absorption coefficients. As implemented, the model is not intended to be general or extensible to other gas species and was never presented as such. Adding additional species would indeed be a non-trivial undertaking, but that is neither the goal nor purpose of RadLib as a software package, which presents and validates models rather than evaluating them. **

9. Documentation:
    1.	Generating the documentation directly via doxygen v. 1.8.20 gives numerous warnings about obsolete tags and then fails because it is looking for a relative path “../../docs/doxygen” that doesn’t exist. This should be fixed.
    1.	Generating the doxygen documentation via the build system target (“make doxygen”) does seem to work fine.

@@

1. RADCAL data were used to generate the PlanckMean absorption coefficient. Based on the reviewer's experience, the PlanckMean absorption coefficients can have meaningful differences when generated from RADCAL and from HiTEMP2010. It is recommended a more recent database being used for the library.

2. Page 10, missing information on [INSERT RUNTIME COMPARISONS HERE].

	**This was a placeholder statement that should not have been included in the document. It has been removed.**

3. In Figure 2, the RCSLM seems to consistently perform better than WSGG, except in Example S2. In addition, WSGG performs worse than PM in Example S4. Can the authors explain why?

4. Equation 6 appears to be inaccurate.

	** Upon review, the authors find no errors in Equation 6, which matches Equation 20.65 of the general band model (of which the WSGG model is a subset) derivation in [2] and appears to be consistent with the other equations near it. **

5. Typo: page 4, line 49, "relating them" -> "relating the"

### 2021-03-24 manuscript/manuscript.tex
Updates to response to reviewers document and associated changes to the paper text.

\begin{small}
\noindent
{\em Program Title:} RadLib                                         \\
{\em CPC Library link to program files:} (to be added by Technical Editor) \\
{\em Developer's repository link:} https://github.com/BYUignite/radlib \\
{\em Code Ocean capsule:} https://codeocean.com/capsule/0997975/tree \\
{\em Licensing provisions(please choose one):} MIT \\
{\em Programming language:} C++, Python                                   \\
{\em Nature of problem(approx. 50-250 words):} \\
Radiation property models can be difficult to understand and implement, often requiring a high degree of specialization. This presents a problem for practitioners who wish to use/implement radiation models for which such models consitute only a small but important part of a larger simulation. Turbulent combustion simulations are one such example. Often, rudimentary assumptions are made and this can negatively impact results. \\
{\em Solution method(approx. 50-250 words):}\\
We provide a well-written, documented, and validated radiation property library that is open source, and uses modern versioning build tools. C++ and Python interfaces are provided along with several examples demonstrating usage of the model. The library can be extended to other models and interfaces. \\ 
{\em Additional comments including restrictions and unusual features (approx. 50-250 words):}\\
The library is intended to be used in Linux-like terminal applications.  
   \\

@@

Modeling radiative heat transfer for CFD is complex and challenging. Combustion simulations, for example, which require many physics models, are often complicated by the need for accurate radiation modeling, which can be difficult to implement and computationally expensive relative to simulation cost. This can lead to simulations that use oversimplified radiation models, but in cases where radiation is significant relative to the other physical phenomena, oversimplified models can negatively impact the accuracy of simulation results. In some cases, radiative gains or losses are small compared to other energy sources or heat transfer modes and radiation can be safely neglected, but many practical engineering systems do not allow this. For example, simple jet flames demonstrate both possibilities: early in a flame's development, turbulent mixing typically dominates heat transfer and dictates flame behavior; late-stage flame phenomena like soot behavior and flame sheet breakthrough, however, increasingly depend on the magnitude of radiative heat losses and are difficult to simulate accurately without a robust radiation model. These problems motivated development of the RadLib library, which provides a framework and uniform interface for radiation property models and implements three powerful models in C++ and Python, giving researchers more convenient access to advanced radiation property models and a consistent framework for further model development.

@@

RadLib makes researchers' work easier by consolidating various approaches into a modular library of interchangeable, prevalidated radiation models. Additionally, RadLib's modular framework is designed to easily accommodate new models as well, allowing researchers to compare new or existing models with very little overhead. Sometimes, it is not clear which radiation model may be the best fit for a particular simulation. When simulations are especially complex or computationally expensive, researchers may have to extrapolate from theory and literature to choose an appropriate radiation model rather than testing for their specific case. RadLib is designed to ease both of these obstacles by facilitating comparison between radiation models through a common interface and providing a practical means of testing various models without the restriction of prespecified geometry or case parameters. With RadLib, researchers can put more of their time and effort into useful results rather than code or model development.

@@

\section{Declaration of Competing Interes} \label{s:coi}

@@

%\bibliographystyle{elsarticle-num} 
%\bibliography{references}

>>>>

\begin{small}
\noindent
{\em Program Title:} RadLib                                         \\
{\em CPC Library link to program files:} (to be added by Technical Editor) \\
{\em Developer's repository link:} https://github.com/BYUignite/radlib \\
{\em Code Ocean capsule:} https://codeocean.com/capsule/0997975/tree \\
{\em Licensing provisions(please choose one):} MIT \\
{\em Programming language:} C++, Python                                   \\
{\em Nature of problem(approx. 50-250 words):} \\
Radiation property models can be difficult to understand and implement, often requiring a high degree of specialization. This presents a problem for practitioners who wish to use/implement radiation models for which such models consitute only a small but important part of a larger simulation. Turbulent combustion simulations are one such example. Often, rudimentary assumptions are made and this can negatively impact results. \\
{\em Solution method(approx. 50-250 words):}\\
We provide an open source radiation property library that uses modern versioning build tools, including C++ and Python interfaces and examples demonstrating usage of the models. The library can be extended to other models and interfaces. \\ 
{\em Additional comments including restrictions and unusual features (approx. 50-250 words):}\\
The library is intended to be used in Linux-like terminal applications.  
   \\

@@

Modeling radiative heat transfer for CFD is complex and challenging. For example, combustion simulations are often complicated by the need for accurate radiation modeling, which can be difficult to implement and computationally expensive relative to overall simulation cost. This can lead to simulations that use oversimplified radiation models, but in cases where radiation is significant relative to the other physical phenomena, oversimplified models can negatively impact the accuracy of simulation results. In some cases, radiative gains or losses are small compared to other energy sources or heat transfer modes and radiation can be safely neglected, but many practical engineering systems do not allow this. For example, simple jet flames demonstrate both possibilities: early in a flame's development, turbulent mixing typically dominates heat transfer and dictates flame behavior; late-stage flame phenomena like soot behavior and flame sheet breakthrough, however, increasingly depend on the magnitude of radiative heat losses and are difficult to simulate accurately without a robust radiation model. These problems motivated development of the RadLib library, which provides a framework and uniform interface for radiation property models and implements three powerful models in C++ and Python, giving researchers more convenient access to advanced radiation property models and a consistent framework for further model development.

@@

RadLib makes researchers' work easier by consolidating various approaches into a modular library of interchangeable, prevalidated radiation models. Additionally, RadLib's modular framework can also accommodate new models, allowing researchers to compare new or existing models with very little overhead. Sometimes, it is not clear which radiation model may be the best fit for a particular simulation. When simulations are especially complex or computationally expensive, researchers may have to extrapolate from theory and literature to choose an appropriate radiation model rather than testing for their specific case. RadLib is designed to ease both of these obstacles by facilitating comparison between radiation models through a common interface and providing a practical means of testing various models without the restriction of prespecified geometry or case parameters. With RadLib, researchers can put more of their time and effort into useful results rather than code or model development.

@@

\section{Declaration of Competing Interests} \label{s:coi}

@@

\bibliographystyle{elsarticle-num} 
%\bibliography{references}
